---
title: "Cap2"
author: "maddi"
date: "`r Sys.Date()`"
output: word_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract

## Introduction

Biodiversity plays a vital role in maintaining the functionality of ecosystems. The results of previous experiments have demonstrated that an increase in species richness can lead to improvements in ecosystem functioning (e.g., increase in biomass due to higher plant species richness [@cardinale2004]). In particular, recent research has sought to elucidate the impact of diversity at different trophic levels on ecosystem functionality, with a particular focus on those related to pollination and plant reproductive success (REF BASCOMPTE JORDANO). It has been demonstrated that an increase in pollinator species diversity can enhance the reproductive success of plants (REF artamendi et al ).

The majority of studies in this field have traditionally focused on biodiversity through the lens of species richness (REF). However, it is equally important to consider the structure of the community. It is insufficient to merely enumerate the number of species; rather, an understanding of the manner in which species are organized and interact within an ecosystem provides a more comprehensive account of its functionality and stability (magrach 2021). Community structure exerts a significant influence on the manner in which the constituent components of an ecosystem contribute to its overall functioning. This underscores the necessity for studies that integrate both species richness and community-level analyses in order to gain a more profound understanding of biodiversity and its role in ecosystem sustainability[@magrach2023].

In particular, we have focused on a static view by pooling data, where if one species interacts with another at one time, it is assumed that this will always happen. This means that we have not investigated how these interactions may vary over time, where different variables may influence them. For instance, CaraDonna and Waser (2020) discovered that pollinator species typically classified as generalists in the overall network were, in reality, significantly more specialized when analyzed on a week-to-week basis. Now is the time to study the dynamics of community patterns over time. HABLAR MAS SOBRE EL TIEMPO, VARIACIONES EN UN MISMO AÑO, PRIMAVERA

DIFERENCIAS ENTRE ESTRUCTURA COMUNIDAD Y A NIVEL DE ESPECIE. macroscala y microescala.

Here, we present an empirical study linking pollinator visitation and plant reproductive success at the community level and species level. We use data collected within two different habitat types (Atlantic mountain grasslands and the understory of Mediterranean woodlands) across two consecutive years in the Mediterranean woodlands and three consecutive years in the mountain grasslands. Our aim is first to understand if macroscale community features (nestedness, functional complementarity) change over small scale time (weekly networks)) and big scale time lapses (seasonally aggregated networks). Second, we ask if microscale features of species (degree of generalization, closeness) vary also through the different time scales. Third, we focus on understanding if adding information on selected interaction network structure indices to previously used simple visitation metrics (e.g., number and diversity of pollinator species) explains better the differences observed in reproductive success.

Specifically, our research aims to answer the following questions at two different levels (site or community level, and species level): (i) at the site level, (1) how dynamic are nestedness and functional complementarity through time in the two climatic zones? do they explain community level reproductive success? ; at the species level (2) does the degree of pollinator generalization change through time? and is it link to species level plant reproductive success? (3) Is plant or pollinator species closeness dynamic through time, and what is its relation to plant species reproductive success?

## Methodology

### Study sites

We collected data in two different study areas and habitat types in Spain: within the vicinity of Doñana National Park in SW Spain and at Gorbea Natural Park, located in N Spain. Doñana features a Mediterranean climate, with warm, dry summers and cool, humid winters. Annual precipitation is 500 mm, and mean temperatures during the study period range from 12.5 to 22.5ºC (REF pizarro2021). Surveys were conducted within fragmented woodlands of stone pines (Pinus pinea) that host a rich understory of flowering shrubs and annual plants (Aparicio 2007). In turn, surveys in Gorbea were conducted within an area characterized by the influence of the Atlantic Ocean, with mild winters and summers and high annual precipitation values, \~ 1,300 mm. Here, mean temperatures during the study period range from 10 to 17.5 ∘ C (Pizarro et al. 2021). Vegetation in the area is composed of beech forests, in conjunction with abundant coniferous plantations. There is also a substantial presence of shrubland and meadows, together with dispersed outrocks, present mainly in the steepest areas of the mountain (Albizu et al. 2002).

Surveys were conducted during two consecutive years (2020 and 2021) in Doñana and three consecutive years (2020, 2021 and 2022) in Gorbea. Within each area we selected five sites within similar elevations (ranging from 50 to 150 m a.s.l. in the case of Doñana and between 800 and 900 m.a.s.l in the case of Gorbea), as well as within similar habitat and soil types, reducing potential confounding factors. Similarity in plant composition between sites was 0.41 ± 0.04 in both areas and across both study years (plant mean Sørensen beta-diversity). The average distance between sites ranged from 3.2 to 24.2 in the case of Doñana and from 1.3 to 4.6 km in Gorbea. In Doñana, one of the sites had to substituted in 2021 due to impossibility of access in the second year of sampling and it was changed to a nearby very similar area.

### Plant pollinator interactions

Within each site, we established one 200-m x 1 m transect. Transects were visited every week (period), when weather conditions allowed, and surveyed three times during each visit, one in the early morning (9.30-10.30), one in the middle of the day (13.30-14.30) and one at the end of the day (16.30-17.30), in order to obtain a complete picture of the community of plants and pollinators present throughout the day. During the whole flowering season (from February to May in the case of Doñana and from March to July in the case of Gorbea), we were able to sample each of the sites between 6 and 9 times depending on the year. Sampling was interrupted for two weeks in 2020 due to the lockdowns derived from the COVID-19 pandemic. The objective was to sample all five sites for one full day each during the week, and then repeat the sampling again the following week in a randomized order, such that each week represented a period with a difference of 4 days in the sampling of each site. This was however not always possible due to lockdowns, restrictions and bad weather, and therefore in a small number of occasions, a period covers a maximum of 10 days difference between the sampling of the first site and that of the last. At each census, along each transect we identified all plant species and recorded all the floral visitors that landed on their flowers and touched the plant´s reproductive parts. Only floral visitors (from now on referred to as pollinators) that could not be identified in the field were captured, stored and identified in the laboratory by an expert entomologist (see acknowledgements). In addition, for each period, we recorded the number of flowers produced by each plant species present in the transect. All surveys were done under similar weather conditions, avoiding windy or rainy days. Total sampling time was XX (435 20-21rako) hours for both areas and years.

### Plant reproductive success

To obtain a measure of pollination function, we estimated the average number of fruits per flower (fruit set) in Doñana´s sites during 2021 and in Gorbea´s sites during 2021 and 2022. For this objective, we focused on the reproductive success for a subset of plant species in the community. Specifically, we focused on abundant species that are also highly self-incompatible, as well as being straightforward to identify, collect and mark. We recorded the reproductive success for different periods and sites. The selected subset includes six species of the Cistaceae family for Doñana: *Cistus salviifolius, C. crispus, C. ladanifer, C. libanotis, Halimium halimifolium,* and *H. calcynum*. We estimated the average number of fruits per flower within 1853 individual flowers throuhout the flowering season of 2021 in Doñana. In the case of Gorbea, we marked and obtained reproductive success data for the following plants throughout the flowering season of 2021 and 2022: *Pedicularis sylvatica, Oxalis acetosella, Primula veris, Viola riviniana, Erysimum gorbeanum, Hepatica nobilis, Vicia pyrenaica, Cardamine pratensis, Lotus corniculatus, Scilla verna, Helleborus viridis, Primula elatior, Ranunculus ficaria, Lathyrus linifolius, Hutchinsia alpina, Hippocrepis comosa, Helianthemum nummularium,* and *Erythronium dens-canis*.

The fruit set was scaled between 0 and 1 by dividing the total number of fruits produced by the number of marked flowers. A value of 1 for fruit set indicates that all marked flowers produced fruit. However, for certain species in Gorbea, such as *Helleborus viridis*, which can produce multiple fruits per flower (up to a maximum of 5), an additional adjustment was made to the fruit set calculation. Specifically, in this case, the total number of fruits was divided by 5 to reflect the potential maximum fruit output per flower.

### Data analyses

#### Sampling completeness

To evaluate the sampling completeness, we estimated the asymptotic number of species of plants, pollinators and interactions present[@chao2009; @chao2020], a non-parametric estimator of species richness for abundance data. This estimator includes non-detected species and allowed us to calculate the proportion detected with our original data. We used Chao 1 asymptotic species richness estimators [@chao2009] and estimated the richness of pollinators, plants and plant-pollinator links accumulated as sampling effort increased up to 100% sampling coverage using package iNEXT [@iNEXT] within the R environment (ref R). We then extracted the values covered by our sampling. We calculated the completeness of the entire network for each year and each study area. Additionally, we estimated completeness for each individual site over different time periods, as well as for each time period across all sites. These different levels of aggregation (by year, site, and time period) were used throughout our analyses.

#### Network analysis

To evaluate differences in network structure between communities, we constructed plant-pollinator interaction networks for each study site (Gorbea and Doñana), sites, and periods or rounds of sampling. We thus obtained 173 interaction networks, representing the number of individuals of different pollinator species recorded visiting each different plant species. For each network, we extracted a series of relevant network metrics at the species and site levels.

**Site-level analysis**

At the site level, information regarding nestedness, pollinator and plant niche complementarity was extracted and total visits conducted on each sampling day were subsequently calculated. Mean plant reproductive success at the community level was used as an index of community functionality. Nestedness is the property where specialist species interact with a subset of the species that generalists interact with (REF BASCOMPTE). Some theoretical studies have found that nested networks are more stable and resilient to perturbations and species extinctions (REF) because nestedness promotes a greater diversity by minimizing competition among species (REF). However, many network attributes vary with network size and complexity (REF), as nestedness can be affected by connectance and network size (REF). We thus used a normalized value of the widely used nestedness metrc NODF based on binary matrices (REF), NODFc (SONG 2017). This normalized value is calculated as NODF~c~ = NODF~n~ / (C \* log(S)), where C is connectance and S is network size. NODF~n~ is calculated as NODF/max(NODF), which is independet of network size and thus comparable across different networks (SONG). To calculate max(NODF) we used a corrected version of the algorithm (SIMMONS) whenever possible.

To calculate niche complementarity, we used a community-level measure defined as the total branch length of a dendogram based on qualitative differences in visitor assemblages between plants (Devoto 2012). All network metrics were calculated using package bipartite[@bipartite] and the corresponding changes recommended by SONG et al 2012. Finally we calculated the total visits performed in each sampling the day in the community.

[@hoeppke2021] maxnoedf indice

**Species-level analysis**

At the species level, we focused on attributes defining the position of a focal plant species within the larger community. (i) Average niche overlap in terms of pollinators between a focal plant species and each of the other plant species in the community (MORISITA). Niche overlap estimates the potential indirect interactions between plant species through shared resources (in this case pollinators) an the potential for increased heterospecific pollen deposition (Arceo_gomez2019). We calculated it as the average overlap in pollinator species visiting a focal plant and each of the other plants in the community using the Morisita overlap index (zhang 2016). (ii) Weighted closeness as a centrality measure as a representation of how close a species is to all other species in the network (REF). Weighted closeness takes into account not only the number of interactions of species but also their strength, meaning that species with many strong interactions with other species will have higher weighted closeness and therefore, a higher potential to impact the network´s dynamics (REF). (iii) Index *d* of specialization of a species which is calculated comparing the observed interactions to an aleatory distribution. Therefore, if a species deviates significantly from what it is expected and interacts just with a subset of the species, it is specialist and will have a high *d* index. Both index, weighted closeness and *d* index of specialization were calculated with the bipartite package in R. Finally, we calculated the total visits per plant species performed in each sampling day.

### Statistical analysis

Having determine each plant species´ role in each network, we wanted to compare the roles of each species between and within sampling periods and years, to test if roles changed over time and if this changes were related to changes in community composition. We also compared how community structure changed over time and through the different sampling sites. Moreover, we wanted to evaluate whether adding information on network structure improves our ability to explain differences in reproductive success, both at the species level and the site level. To evaluate this, we used generalized linear mixed models (GLMMs) respectively. To standardize the continuous variables in our GLMM models, we used the scale() function in R, which transforms the variables by subtracting the mean and dividing by the standard deviation (SD). This transformation ensures that all variables have a mean of 0 and a standard deviation of 1, allowing for the comparability of the model coefficients. We thus transformed the following variables: pollinator species richness, plant species richness, total average visits per site, total visits per plant species and site, nestedness, functional complementarity for plants and pollinators, niche overlap (Morisita), *d* index of specialization and weight closeness as the an index of centrality.

We analyzed the influence of network metrics as explanatory variables on site and species level reproductive success using beta-family distribution models. The beta family is particularly suitable for modeling response variables that are continuous and constrained within the interval (0, 1), as it allows for the modeling of non-normal distributions. In order to evaluate the variations in time and space observed in the different network metrics and to determine whether these variations are influenced by other factors, Gaussian distribution models were employed. All models were performed in parallel for both the Gorbea and Doñana datasets. In both cases, the sampling sites (five sampling sites for each study site) were included as a random factor. From these models, we derived predictive values for subsequent visualisation for each study site and each year, as well as secondary figures also visualised by sampling site.

## Results

Across both our study areas and years, our sampling spanned XXX h of observation. During this time, we recorded a total of 61 and 90 plant species in Doñana and Gorbea, respectively. Out of these, 75.4 % of the plant species found in the different sites in Doñana, and 72.2 % of the plant species present in Gorbea received at least one visit from floral visitors. Within our sampling we recorded 3939 interactions (1690, 956 and 1293 for 2020, 2021 and 2022) involving 155 pollinator species and 65 plant species for Gorbea and 2617 interactions (879 and 1738 for 2020 and 2021 respectively) involving 144 pollinator species and 46 plant species for Doñana (Table 1). We recorded a total of 452 and 646 unique combinations of plants and flower visitors in each of the study areas (Doñana and Gorbea) through the time. These interactions involved 197 and 282 distinct combinations of plants and flower visitors for Doñana in 2020 and 2021 (being 37.61% and 56.42% of them unique for those years) and 282, 208 and 301 distinct interactions for Gorbea in 2020, 2021 and 2022 (being 30.03%, 19.81% and 32.82% of them exclusive for each year).

In Doñana, the most represented order of flower visitors was Hymenoptera, involved in 69.51 % of the interactions observed. Specifically, Apidae represented the most common family within the order, representing 48.34 % of the interactions and interactions involving Apis mellifera, the managed European honeybee, represented 41.77 % of all interactions. Following Hymenoptera, the next most common order was Diptera representing 15.28 % of the interactions, and specifically the family Syrphidae representing 10.39 %. In turn, Coleoptera represented 13.34% of the interactions and Lepidoptera 1.75%. In Gorbea, the most abundant group were Diptera which accounted for 63.03 % of the interactions, and specifically the family Syrphidae represented 52.75 %, followed by Hymenoptera, representing 29.93 %, and specifically the family Apidae representing 19.01 %, but in this case Apis mellifera represented 8.55 % of the interactions and different species in the Bombus genus 9.7 %. Coleoptera and Lepidoptera represented 3.7 % and 3.17 % of interactions.

VIOLIN PLOT ABUNDANCE TAXONOMIC GROUP PER PERIOD FOR EACH SITE AND YEAR

### Sampling completeness

Our sampling completeness analyses revealed that our survey was able to capture high sampling coverage for both plant and pollinator communities in both study years. We detected 81% and 77% of pollinator species in 2020 and 2021, respectively, in the case of Doñana, and 66%, 54%, and 50% in the case of Gorbea in 2020, 2021, and 2022 (Fig 2 and Supplementary Fig. X). Regarding plant species, we detected 95% and 88% for 2020 and 2021, respectively, in Doñana, and 92%, 94%, and 92% for 2020, 2021, and 2022 in Gorbea. In the case of plant-pollinator links, sampling completeness is lower in both study areas (60% and 60% for 2020 and 2021 in the case of Doñana and 44%, 51% and 53% in the case of Gorbea) for the three sampling years.

Nevertheless, these values are subject to alteration when alternative amplitudes or scales are taken into account during the calculation of sampling completeness values. For example, if sampling completeness is calculated aggregating data across sites for each of the different periods, we obtain highly variable coverage values ranging from 47.03% to 75.82% in the case of pollinators in Doñana and from 21.52 to 91.98% in the case of Gorbea in 2020 (Supplementary Fig. X and Table X). Similarly, in the case of plant species, sampling completeness is more variable when period is considered as sampling breadth, with between 65.48 to 98.47% of plant species being identified in the case of Doñana and between 36.32% and 95.36% in the case of Gorbea in 2020. In the case of links, these values ranged from 7.25% to 64.73% in the case of Doñana and between 36.86 and 81.3 % in Gorbea. Aggregating data through time for each site, showed values ranging between 37% and 84.58% for pollinator species detected in Doñana in 2020, while in Gorbea's different sites, the values ranged between 42.43% and 65.51%. The proportion of plant species detected in Doñana ranged from 74.37% to 94.17%, while in Gorbea it ranged from 57.7% to 96%. In the case of links, the corresponding values were 41.64% to 65.23% in Doñana and 22.99% to 42.98% in Gorbea.

An inspection of interactions shows that 37.61% and 56.42% of the interactions observed in Doñana in 2020 and 2021 are unique to that year, i.e., they are not observed the other year (Table 1). In turn, in Gorbea these proportions stand at 30.03% , 19.81% and 32.82% respectively for 2020, 2021 and 2022. In turn, pollinator species that are uniquely found in one of the study years represent 25.69% and 50% of all pollinator species in 2020 and 2021 respectively in the case of Doñana, and 27.10% ,11.61% and 23.23% in the case of Gorbea. For plant species involved in interactions, values of species uniquely found to interact one of the study years stand at 34.78% and 32.61% in 2020 and 2021 for Doñana and 9.23%, 4.62% and 16.93% for Gorbea.

### Community network structure 

With respect to the temporal dynamics of plant-pollinator networks´ nestedness, markedly divergent patterns emerge across the sampling seasons within the two study areas, as illustrated in Fig. X.A-B. However, when focusing on the patterns observed within each study site across different sampling years, we could appreciate similarities, where the network structure of Gorbea displayed a greater degree of fluctuation in its nestedness. In Gorbea, the data indicated a decrease in nestedness at the start of the sampling season, which then increases midway through and stabilises at higher values. This corresponds to a greater degree of nestedness within the network. In Doñana, we observed that the level of nestedness tends to increase throughout the sampling period, reaching a peak in the middle of the sampling period. However, as the sampling period progresses with spring, the level of nestedness begins to decrease. Concurrently, our findings revealed that the interaction networks of the Gorbea sampling areas exhibited considerably greater variability, both between and within the same sampling site across different years, in comparison to the Doñana sampling areas (Supplementary Fig. XA-B). In contrast, the opposite occurred in the Doñana sampling areas, where a consistent pattern is observed in all five sites in both years throughout the sampling period.

With regard to the functional complementarity of pollinators and plants, the same temporal patterns were observed for both of them. As was observed in the Gorbea nesting network, the functional complementarity of both pollinators and plants exhibited greater fluctuation in Gorbea, with the same pattern being observed for all three years of sampling. The highest values of functional complementarity were detected in the middle of the sampling season.

### Community level reproductive success

### Species complementarity through time and space

```{r species_presence}
load("scripts/presence_plot.RData")
presence_plot
```

### Species level reproductive success

```{r cars}
summary(cars)
```

## Discussion

```{r pressure, echo=FALSE}
plot(pressure)
```

## References
